---
title: "Description packages and functions used for tree biomass calculations"
author: "MarteF"
format: html
editor: visual
---

```{r}
# install.packages("sitreeE")
# install.packages("sitree")
# install.packages("gt")

library(sitree)
library(sitreeE)
library(gt)
```

## Introduction to packages used

For this calculation we have used the packages "sitree" and "sitreeE" developed by Clara Anton Fernandez.

Citation: Anton Fernandez C (2021). \_sitreeE: Sitree Extensions\_. R package version 0.0-8, \<https://CRAN.R-project.org/package=sitreeE\>.

In these package we find functions to calculate above- and belowground biomass of certain nordic trees like spruce, pine and birch, based on known biomass equations.

The **sitree** package provides a framework to implement Single Tree forest growth models in a fast and memory efficient way. It keep tracks of all alive, dead, and removed trees in a robust and efficient way. **SiTree** is designed to run single tree simulations where trees can be defined by two time-dependent variables (such as diameter (or basal area), and height), and on time-independent variable, such as tree species. **SiTree** simulates birth, growth, and death of trees as well as management. Functions can also be defined that affect characteristics of the stand (external modifiers), such as climate change, or fertilization.

SitreeE provides extensions for package 'sitree' for allometric variables, growth, mortality, recruitment, management, tree removal and external modifiers functions.

It supports the simulation of different management interventions and their effects on forest growth and development.

Two types of input are required by **SiTree**: tree level and stand level. Tree level information is passed in `tree.df`, while stand level information is passed in `stand.df`.

`tree.df` should be a data frame with four columns named `plot.id`, `treeid`, `dbh`, `height`, and `tree.sp`, which correspond to a stand/plot ID, a tree ID, diameter, height, and tree species.

Plot and stand data is passed in `stand.df`, which should be a data frame or a list, with at least a column or element named `plot.id` which should contain all the plot IDs present in `tree.df`. Typical information provided in `stand.df` are plot size, elevation, site index, plot coordinates, distance to road, temperature or precipitation.

e.g.

```{r}

# Example data
data <- data.frame(
  plot.id = c(91, 91, 91, 91, 91, 91),
  treeid = c(108286, 137120, 108287, 108268, 137124, 108271),
  dbh = c(149, 133, 55, 235, 187, 162),
  height = c(118, 107, 52, 151, 137, 124),
  tree.sp = c(53, 53, 53, 49, 53, 49)
)

# Create a formatted table
data %>%
  gt() %>%
  tab_header(
    title = "Tree Data Table",
    subtitle = "Example of tree measurements"
  )
```

### Marklund’s biomass equations

biomass.spruce.M1988(dbh.cm, H.m)

biomass.pine.M1988(dbh.cm, H.m)

biomass.birch.M1988(dbh.cm, H.m)

Arguments dbh.cm is a vector with the dbh (diameter at breast height) of the trees, in cm.

H.m is a vector with the heights of the trees, in meters.

It returns a data.frame with the following biomass components in kg: living.branches, dead.branches, stem.wood, stump.roots, bark, usoil, rot1, rot2, and foliage.

References Marklund, L. G. 1988. “Biomassafunktioner för tall, gran och björk i Sverige \[Biomass functions for pine, spruce and birch in Sweden\].” Report 45. Umeå, Sweden: Swedish University of Agricultural Sciences. Department of Forest Survey. Petersson, Hans, and Göran Ståhl. 2006. “Functions for Below-Ground Biomass of Pinus Sylvestris, Picea Abies, Betula Pendula and Betula Pubescens in Sweden.” Scandinavian Journal of Forest Research 21 (S7): 84–93.

### biomass functions for birch for Norway from Smith et al (2014, 2016)

biomass.birch.S2014

Returns a data.frame with the following biomass components: biomass.total.kg, biomass.aboveground.kg.S2014, biomass.belowground.kg.S2014, biomass.belowground.kg, biomass.aboveground.kg, living.branches, dead.branches, stem.wood, stump.roots, bark, usoil, rot1, rot2, foliage)

It uses stump.roots from biomass.birch.M1988 to calculate aboveground biomass (in kg) and belowground biomass (in kg). Aboveground biomass is calculated as Smith’s aboveground biomass (2014) + stump.roots. Belowground biomass is calcualted as Smith’s belowground biomass (2016) - stump.roots.

References Smith, Aaron, Aksel Granhus, and Rasmus Astrup. 2016. “Functions for Estimating Belowground and Whole Tree Biomass of Birch in Norway.” Scandinavian Journal of Forest Research 31 (6): 568–82. <https://doi.org/10.1080/02827581.2016.1141232> Smith, Aaron, Aksel Granhus, Rasmus Astrup, Ole Martin Bollandsås, and Hans Petersson. 2014. “Functions for Estimating Aboveground Biomass of Birch in Norway.” Scandinavian Journal of Forest Research 29 (6): 565–78. <http://dx.doi.org/10.1080/02827581.2014.951389>

### The Norwegian biomass equation

biomass.norway.sitree(dbh.mm, height.dm, tree.sp)

Arguments:

dbh.mm - Diameter at breast height in mm.

height.dm - Tree height in dm.

tree.sp Tree species according to the Norwegian NFI.

It will use biomass.spruce.M1988 when tree.sp is 1, 2, 3, 21, or 29, biomass.pine.M1988 when tree.sp is 10, 11, or 20, and biomass.birch.S2014 otherwise.

## Tree species codes 

-   **Norway Spruce (*Picea abies*)**: Code 1

-   **Scots Pine (*Pinus sylvestris*)**: Code 10

-   **Birch species (*Betula spp.*)**: Code 30

-   

## Example

```{r}
bio.nor <- biomass.norway.sitree (dbh.mm = c(50, 60, 100 ), height.dm = c(40, 60,
                                                                          80),
                                  tree.sp = c(1, 10, 30))




```

```{r}
data(tr)
```
